---
title: "R/RStudio/RMarkdown Recap"
author: "Heili Lowman"
date: "9/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Demo Dataset

In the following RMarkdown, I will be using the `lynx` dataset from `data()` in base R to practice importing, examining, tidying, editing, and visualizing a dataset.

Remember - to insert a new code chunk either click on **Insert** > **R** at the top of the screen or use one of the shortcuts: 

* Mac - **cmd + opt + I**
* PC - **ctrl + alt + I**

### Data & Packages

```{r packages, message = FALSE}

# Load necessary packages.
library(tidyverse)

# Add "message = FALSE" to the header to suppress messages from the knit document.

```


```{r data}

# Load dataset.
lynx_data <- lynx

# If your data is external, save it to your R project, then import using read_csv().
# dataset <- read_csv("my_data.csv")

# Examine data. Then comment out, so it doesn't run in your script.
# View(lynx_data)

```


### Data Tidying & Manipulation

```{r tidy}

# Examine dataset structure. Also commenting out so it doesn't run.
# str(lynx_data)

# My dataset is a time series, which can be a helpful format for some analyses, but I don't want it that way for this purpose.

# So, this makes it a normal dataframe/tibble.
# install.packages("timetk")
library(timetk)
lynx_df <- tk_tbl(lynx_data)

# Rename columns.
lynx_df <- lynx_df %>%
  rename(Year = index, Lynx = value)

```


### Data Visualization

```{r plot, message = FALSE}

# Create a basic scatterplot of the data.
lynx_fig <- ggplot(lynx_df, aes(x = Year, y = Lynx)) + # basic data and axes input
  geom_point(size = 3, alpha = 0.75) + # calls for a scatterplot
  geom_smooth(method = lm, color = "aquamarine2") + # adds a linear model with 95% confidence intervals
  labs(x = "Year",
    y = "Number of Lynx Trapped",
    title = "Lynx R Dataset",
    subtitle = "Tidy Tuesday Demo, September 1",
    caption = "Data Source: Brockwell and Davis 1991") + # includes labels
  theme_classic() + # uses a built in theme
  scale_x_continuous(breaks = seq(1820,1940,10)) ## customizes x axes tick marks

lynx_fig

```

*End of RMarkdown script.*